Expanding the scope of experimental archaeology using the Perception-Process-Product conceptual framework

### 2023-05-23 manuscript/manuscript.Rmd
minor revision.
added new references.

Ethological approaches has been first systematically developed and applied in the archaeological research by Haidle [@haidleHowThinkSimple2009; @haidle2010; @lombard2012], known as cognigram, essentially representing an abstracting process of a series of behavioral sequences achieving a similar goal. This approach is a power and elegant yet limited by the curse of expertise [@hinds1999]. Like chaine operatoire, it cannot handles variation very well. To some extent it describes the minimal steps to achieve a goal from the perspective of reverse engineering and assume clear causal thinking between each steps. Novices has a different sets of perception on the causal structure of how certain behaviors will modify the raw materials, leading to over-imitation. Here we used the ethogram, or the action grammar, developed by [@stout2021] as an example. Other coding scheme also exist such as [@mahaney2014], or rotation analysis [@muller2023], or [@cueva-temprana2019].

>>>>

Ethological approaches has been first systematically developed and applied in the archaeological research by Haidle [@haidleHowThinkSimple2009; @haidle2010; @lombard2012; @haidle2023], known as cognigram, essentially representing an abstracting process of a series of behavioral sequences achieving a similar goal. This approach is a power and elegant yet limited by the curse of expertise [@hinds1999]. Like chaine operatoire, it cannot handles variation very well. To some extent it describes the minimal steps to achieve a goal from the perspective of reverse engineering and assume clear causal thinking between each steps. Novices has a different sets of perception on the causal structure of how certain behaviors will modify the raw materials, leading to over-imitation. Here we used the ethogram, or the action grammar, developed by [@stout2021] as an example. Other coding scheme also exist such as [@mahaney2014], or rotation analysis [@muller2023], or [@cueva-temprana2019].
